---
title: "p8105_hw1_xs2330"
author: "Xiao Shi"
date: "September 18, 2018"
output: github_document
---
**Set up and set seed to ensure reproducibility**
```{r setup}
knitr::opts_chunk$set(echo = TRUE)
set.seed(8105)
```
## Problem 1 ##
**Creating a data frame comprised with 4 variables**
```{r construction, echo=TRUE}
library(tidyverse)
## set seed to ensure reproducibility
df_problem1 = tibble(
  vec_unif = runif(10, min = 0, max = 5),
  vec_logical = c(vec_unif > 2),
  vec_char = c("My", "name", "is", "Xiao", "Shi", "and", "I", "am", "a", "sailor"),
  vec_factor = factor(c("male", "male", "female", "male","male", "male", "male", "female","male", "male"))
)
```

**Taking the mean of each variable**
```{r mean, warning=FALSE}
df_mean = sapply(df_problem1, mean)
```
The num variable and logical variable got the mean successfully
Warning messages generated for character variable and factor variable, returning NA

**Convert variables that are not numeric to numeric**
```{r non-numeric to numeric, eval=FALSE, warning=FALSE}
vec_logical_converted = as.numeric(df_problem1$vec_logical)
vec_char_converted = as.numeric(df_problem1$vec_char)
vec_factor_converted = as.numeric(df_problem1$vec_factor)
```
Warning message generated when converting character variable to numeric format

**Convert variables that are not numeric to numeric with an extra step**
```{r non-numeric to numeric extra, eval=FALSE, warning=FALSE}
vec_char_converted_extra = as.numeric(as.factor(df_problem1$vec_char))
vec_factor_converted_extra = as.numeric(as.character(df_problem1$vec_factor))
```
Warning meesage generated when converting with an extra step. This time what originally was a character variable converted successfully into numeric through converting to factor first. However, what originally was a factor got converted into character variable before converting to numeric and thus resulted in the warning message.


## Problem 2 ##
**Creating a data frame comprised with variables x and y**
```{r data construction p2}
n = 1000
plot_df = tibble(
  x = rnorm(n = 1000, mean =0, sd = 1),
  y = rnorm(n = 1000, mean =0, sd = 1),
  logical_xy = c(x + y > 0),
  logical_to_num_xy = as.numeric(logical_xy),
  num_to_factor_xy = as.factor(logical_to_num_xy)
)
```
* the size of the dataset is `r n`
* the mean of x is `r mean(plot_df$x)`, and the median of x is `r median(plot_df$x)`
* the proportion of cases for which the logical vector is TRUE is `r mean(plot_df$logical_xy == TRUE)`

**Making ggplot of x and y**
```{r plotting plot_df}
ggplot(plot_df, aes(x = x, y = y, color = plot_df$logical_xy)) + geom_point()
ggsave("scatter_plot_x+y>1.pdf", height = 4, width = 6)
ggplot(plot_df, aes(x = x, y = y, color = plot_df$logical_to_num_xy)) + geom_point()
ggplot(plot_df, aes(x = x, y = y, color = plot_df$num_to_factor_xy)) + geom_point()
```